Difference-in-Differences with Spatial Spillovers

### 2021-03-22 latex/slides_short/Spillover_slides_short.tex
Update slides for Westerns

\begin{frame}{Spatial Spillovers}
    Researchers aim to estimate the average treatment effect on the treated: 
    \[
        \tau \equiv \mathbb{E} \left[ Y_{i1}(1) - Y_{i1}(0) \ \vert \ D_{i} = 1 \right]
    \]

    \vspace{5mm}
    Spillover effects are when effect of treatment extend over boundaries such as states, counties, etc.

    \begin{itemize}
        \item e.g. large employer opening/closing in a county have positive employment effects on nearby counties
    \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Contribution}

    {\footnotesize

    \underline{Difference-in-Differences Estimation with Spillovers}

    \begin{itemize}
        \item I generalize the work by deriving the bias in terms of potential outcomes and show how to estimate treatment effects by removing spillover effects
    \end{itemize}

    \pause
    \underline{Potential Outcome Framework for Spillovers}

    \begin{citecolor}
        [\citet{Miguel_Kremer_2004}, \citet{Vazquez-Bare_2019}, \citet{Angelucci_DiMaro_2016}, \citet{Angrist_2014}]        
    \end{citecolor}

    \begin{itemize}
        \item I focus on a spatial setting whereas these papers consider spillovers in network clusters
    \end{itemize}

    \pause
    \underline{Place-based Policies}

    \begin{citecolor}
        [\citet{Kline_Moretti_2014b}, \citet{Kline_Moretti_2014a}, \citet{Busso_Gregory_Kline_2013}, and \citet{Neumark_Kolko_2010}]
    \end{citecolor}

    \begin{itemize}
        \item I highlight the need to control for general equilibrium effects to properly estimate local effects of policies
    \end{itemize}
    }
\end{frame}

@@

\begin{frame}{Potential Outcomes Framework}
    For exposition, I will label units as counties. Assume all treatment occurs at the same time (2-periods or pre-post averages).\footnote{I extend this into an event study framework in the paper, but the intuition is the same as in the $2 \times 2$ setting.}

    \begin{itemize}
        \item $Y_{it}(D_i, \textcolor{blue}{h(\vec{D}, i)})$ is the potential outcome of county $i \in \{ 1, \dots, N \}$ at time $t$ with treatment status $D_i \in \{0, 1\}$.

        \pause
        \item $\vec{D} \in \{0,1\}^N$ is the vector of all units treatments.

        \pause
        \item The function $\textcolor{blue}{h(\vec{D}, i)}$ maps the entire treatment vector into an `exposure mapping' which can be a scalar or a vector.

        \pause
        \item No exposure is when $\textcolor{blue}{h(\vec{D}, i)} = \vec{0}$.
    \end{itemize}
\end{frame}

% Grey out
% \setbeamercovered{transparent}

@@

    Examples of $h_i(\vec{D})$:

    \begin{itemize}
        \item \textbf{Treatment within $x$ miles:}

@@

    \end{itemize}
\end{frame}

    Examples of $h_i(\vec{D})$:

    \begin{itemize}
        \item \textbf{Treatment within $x$ miles:}

        \vspace{2.5mm}
        \item \textbf{Number of Treated within $x$ miles:}

@@

    \end{itemize}
\end{frame}

@@

\begin{frame}{Estimand of Interest}
    Estimand of Interest: \[ 
        \textcolor{green}{\tau_{\text{direct}}} \equiv \textcolor{green}{\mathbb{E}\left[ Y_{i,1}(1, 0) - Y_{i,1}(0, 0) \mid D_i = 1\right]}
    \]

    \vspace{10mm}
    This is the direct effect in the absense of exposure to spillovers.
\end{frame}

\begin{frame}{Parallel Trends}
    I assume a modified version of the parallel counterfactual trends assumption: 
    \begin{align*}
        &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 1 \big] \\
        = \ &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 0 \big],
    \end{align*}

    \vspace{5mm}
    In the complete absence of treatment (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status

\begin{frame}{What does Diff-in-Diff identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the diff-in-diff estimate as follows:

@@

\begin{frame}{Biased Estimate for $\tau_{\text{direct}}$}
    \[ 
        \mathbb{E}[\hat{\tau}_{\text{diff-in-diff}}] - \textcolor{green}{\tau_{\text{direct}}} = \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}    
    \]
\end{frame}

% ------------------------------------------------------------------------------
\section{Estimation}
% ------------------------------------------------------------------------------

\begin{frame}{Problems with "Dropping Bad Controls"}
    It is common in empirical applications to drop control units near the treated units when estimating the direct effect of treatment.

    This is not recommended for two reasons:

    \begin{itemize}
        \item[1.] If not all control units that experience spillover effects are removed, then bias remains

        \item[2.] There is a second source of bias, the spillover effects on treated units, that still remains
    \end{itemize}

@@

\begin{frame}{Robustness to Misspecification}
    Generate data using the same data-generating process as before but with different spillover functions:

    \[ 
        y_{\it} = \mu_t + \mu_i + 2 D_{it} + \beta_{\text{spill,control}} * (1-D_{it}) h(\vec{D}, i) + \varepsilon_{it}
    \]

    Then, I estimate each data-generating process using (potentially) misspecified $\tilde{h}(\vec{D}, i)$ and report the average estimate bias.

@@

\begin{frame}{Results}
    \textbf{I find that an indicator for being Within $x$ miles from treated area will remove all bias so long as the indicator contains all the affected units.}

@@

    I repeat the same exercise and estimate the spillover effects for each control unit, $\hat{\beta}_{\text{spill, control}} * \tilde{h}(\vec{D}, i)$.

    Then calculate \[ 
        1 - \frac{
                \overbrace{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i) - \hat{\beta}_{\text{spill, control}} \tilde{h}(\vec{D}, i))^2}^{\text{Mean Square Prediction Error}}
            }{
                \underbrace{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i))^2}_{\text{Normalization}}
            }    
    \]

    This gives the proportion of spillovers explained by $\tilde{h}(\vec{D}, i)$

\begin{frame}{Results}
    \textbf{Rings perform best at estimating spillover effects.}
\end{frame}

>>>>

\begin{frame}{Spatial Spillovers}
    Researchers aim to estimate the \textbf{average treatment effect on the treated}: 
    \[
        \tau \equiv \mathbb{E} \left[ Y_{i1}(1) - Y_{i1}(0) \ \vert \ D_{i} = 1 \right]
    \]

    Estimation is complicated by \textbf{Spillover Effects}

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries (states, counties, etc.)

    \begin{itemize}
        \item E.g. a large employer opening/closing in a \textbf{treated} county have positive employment effects on \textit{nearby counties}

        \item Having nearby counties with factories raises wages and reduces effect of \textbf{treated} counties
    \end{itemize}
\end{frame}

@@

\begin{frame}{Econometric Contribution}

    I formalize spillovers into a potential outcomes framework:

    \begin{itemize}
        \vspace{2.5mm}
        \item I decompose the difference-in-differences estimator into three parts: Direct Effect of Treatment, Spillover onto Treated Units, Spillover onto Control Units

        \vspace{2.5mm}
        \item Show that an indicator for being close to treated units remove \textit{all bias} so long as the indicator contains all units affected by spillovers

        \vspace{2.5mm}
        \item `Rings' are able to estimate spillover effects non-parametrically while still removing all bias
    \end{itemize}

@@

@@

\begin{frame}{Potential Outcomes Framework}

    $Y_{it}(D_i, \textcolor{blue}{h(\vec{D}, i)})$ is the potential outcome of county $i \in \{ 1, \dots, N \}$ at time $t$ with treatment status $D_i \in \{0, 1\}$.

    The function $\textcolor{blue}{h(\vec{D}, i)}$ maps the entire treatment vector into an `exposure mapping'
\end{frame}

\begin{frame}{Examples of $h_i(\vec{D})$}
    \textbf{Treatment within $x$ miles:}

        \item Spillovers are non-additive
    \end{itemize}
\end{frame}

@@

    \textbf{Treatment within $x$ miles:}

@@

    \vspace{2.5mm}
    \textbf{Number of Treated within $x$ miles:}

@@

@@

\begin{frame}{What does Diff-in-Diff identify?}
    With the parallel trends assumption and a random sample, I decompose the difference-in-differences estimate as follows:

@@

@@

\begin{frame}{Robustness to Misspecification}
    I find that an indicator for being Within $x$ miles from treated area interacted with treatment status will remove \textbf{all bias} so long as the indicator contains all the affected units.

    \begin{itemize}
        \item Indicator will estimate the average spillover effect on treated and control units and remove these from estimate, $\hat{\tau}$
    \end{itemize}
\end{frame}

    A set of concentring `rings' around treatment perform best for estimating spillover effects
\end{frame}

\begin{frame}{Benefits of Rings}

    \begin{itemize}
        \item Still remove all bias from the treatment effect estimate

        \item Can trace out how spillovers spread over distance

        \item If spillovers are additive in the number of nearby treated units, then an additive version of rings should be used (but this loses the bias-removal property)
    \end{itemize}

@@

\section{Application in Urban Economics}

\begin{frame}{Urban Economics}

    I apply framework to place-based analyses in Urban Economics

    \begin{itemize}
        \item Revisit \begin{citecolor}\citet{Kline_Moretti_2014a}\end{citecolor} analysis of the Tennessee Valley Authority

        \begin{itemize}
            \vspace{2.5mm}
            \item The local effect estimate is contaminated by spillover effects to neighboring counties \begin{citecolor}\citep{Kline_Moretti_2014b}\end{citecolor}

            \vspace{2.5mm}
            \item Large scale manufacturing investment creates an `urban shadow' \begin{citecolor}\citep{Cuberes_Desmet_Rappaport_2021,Fujita_Krugman_Venables_2001}\end{citecolor}
        \end{itemize}

        \vspace{2.5mm}
        \item Discuss how framework can reconcile conflicting findings on effect of federal Empowerment Zones

        \begin{itemize}
            \vspace{2.5mm}
            \item Identification Strategy of using far-away rejected applicants \begin{citecolor}\citep{Busso_Gregory_Kline_2013} \end{citecolor} vs. census tracts within 1000 feet of empowerment zone \begin{citecolor}\citep{Neumark_Kolko_2010}\end{citecolor}
        \end{itemize}
    \end{itemize}

@@
